---
title: Tidyverse Join Function Part-2
author: ~
date: '2020-01-19'
slug: tidyverse-join-function-part-2
categories: [R, Tidyverse]
tags: []
subtitle: ''
summary: 'Introduction for Join function in Tidyverse package Part2'
authors: [Ye Tao]
lastmod: '2020-01-19T23:32:49-08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---



In the [Part1 Introducton for Join function family in tidyverse family](https://yetao7977.netlify.com/post/test/), we explored some useful functions by exploring some toy examples. In this post, we go further with some slightly more complicated dataset. By doing so helps us have a deeper unstanding how join family works.

First part will just go over the function a liitle bit with a different dataset, and second part will practice in the more complicated datasets.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## The Data

```{r}
superheroes <- tibble::tribble(
       ~name, ~alignment,  ~gender,          ~publisher,
   "Magneto",      "bad",   "male",            "Marvel",
     "Storm",     "good", "female",            "Marvel",
  "Mystique",      "bad", "female",            "Marvel",
    "Batman",     "good",   "male",                "DC",
     "Joker",      "bad",   "male",                "DC",
  "Catwoman",      "bad", "female",                "DC",
   "Hellboy",     "good",   "male", "Dark Horse Comics"
  )

publishers <- tibble::tribble(
  ~publisher, ~yr_founded,
        "DC",       1934L,
    "Marvel",       1939L,
     "Image",       1992L
  )

superheroes
publishers
```

## Inner Join
As a reminder, `inner_join(x, y)` returns all matched rows and columns between `x` and `y` tables. Notice that rows from `publishers` will duplicate themselves for matching the rows from `superheros` in this case.

```{r}
superheroes %>%
    inner_join(publishers, by = "publisher")
```


## Left Join

`left_join(x, y)`: Return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.

```{r}
superheroes
publishers


superheroes %>%
    left_join(publishers, by = "publisher")
```



## Semi join
semi_join(x, y): Return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join.

Recall that `superheros` is the baseline table

```{r}
superheroes %>%
    semi_join(publishers, by = "publisher")
```


## Full Join
full_join(x, y): Return all rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing. This is a mutating join.

```{r}
superheroes %>%
    full_join(publishers, by = "publisher")
```


So far, all the logic basically follows what we have talked in the [previous post](https://yetao7977.netlify.com/post/test/)




## Question

- What if column names in `table A` and `table B` are the same. How we gonna do for this situation.
- what if the colname we wanna join has different names in two tables.


Let's check out the different datatset
```{r}
parts = readr::read_rds(url("https://assets.datacamp.com/production/repositories/5284/datasets/cb649926d41ce73490a9bb710e1501a273061723/parts.rds"))

parts_categories <- readr::read_rds(url("https://assets.datacamp.com/production/repositories/5284/datasets/30fc459770c89e46cce9cce99752ca95fb1d06fe/part_categories.rds"))
```

## glimpse the data

Here we notice that both `table parts` and `table parts_categories` have col name called `name`
```{r}
parts %>% head

parts_categories %>% head
```

Let's pratcice. In this case we are told that `part_cat_id and` from `parts` and  `id` from `parts_categories` are the two columns we want to join. In this case we need to specify which two colnames we wanna join.

```{r}
parts %>%
    inner_join(parts_categories, by = c("part_cat_id" = "id")) %>% head # we need double quote in colname
```

we find that `inner join function` aissgned two different names to the same colname from two table: `name.x` and `name.y`. we can specify the suffox in the name by doing so.


```{r}
parts %>%
    inner_join(parts_categories, by = c("part_cat_id" = "id"),
               suffix = c("_parts", "_part_categoroes")) %>% head
```

Notice that the new table above keeps `part_cat_id` as column name since `parts table` works as the baseline table.
