---
title: Data Tidying with Tidyverse in R
author: ~
date: '2020-01-22'
slug: data-tidying-with-tidyverse-in-r
categories: [R, Tidyverse]
tags: []
subtitle: ''
summary: 'Data tidying with pivot_long and pivot_wide in R'
authors: [Ye Tao]
lastmod: '2020-01-22T19:32:26-08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


## Tidy Data

There are three interrelated rules which make a dataset tidy

- Each variable must have its own column (long format).
- Each observation must have its own row (wide format).
- Each value must have its own cell.

Using the datasets from R for Data Science to show that the same data could be organised in different ways.



```{r}
# make sure you have tidyr 1.0
library(tidyr)
```

`pivot_longer` is the replacement for `gather()` and `pivot_wider()` is the replacement for `spread()`.
Both are designed to be simpler and can handle more cases than gather and spread.

### Pivot longer

```{r}
relig_income %>%
  pivot_longer(-religion, names_to = "income", values_to = "count")
```


```{r}
billboard %>%
  pivot_longer(
    cols = starts_with("wk"),
    names_to = "week",
    names_prefix = "wk",
    names_ptypes = list(week = integer()),
    values_to = "rank",
    values_drop_na = TRUE,
  )
```

### Pivot wider

```{r}
fish_encounters %>% pivot_wider(
  names_from = station,
  values_from = seen,
  values_fill = list(seen = 0)
)
```

```{r}
us_rent_income %>%
  pivot_wider(names_from = variable, values_from = c(estimate, moe))
```


