---
title: General Introduction of Tidyverse Series2
author: ~
date: '2020-01-22'
slug: general-introduction-of-tidyverse-series2
categories: [R, Tidyverse]
tags: []
subtitle: ''
summary: 'Introduction of Tiduverse package part2'
authors: [Ye Tao]
lastmod: '2020-01-22T18:38:24-08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


# **Tidyverse Introduction Part 2**

This is part 2 of the Tidyverse Series; In this post, we continue to explore some extremely useful tools for data analysis

```{r, warning=FALSE, message=FALSE}

library(tidyverse)
library(nycflights13) # Please intstall this packages which contains some interating dataset.
flight_subset <- 
    flights %>% sample_n(5)


flight_subset
```


## **Mutate** : Create new variable

We can use `mutate` to create new variable, and then add them to the end of the origain data set

Recall that `dep_delay`, `arr_delay:` Departure and arrival delays, in minutes. Negative times represent early departures/arrivals.
```{r}
flight_subset %>% 
    mutate(gain = arr_delay-dep_delay, # minutes for getting ealier or later
           speed = distance / air_time * 60, # in miles/hrs
            gain_per_hour = gain / (air_time/60)) # we can also refer the variable we just created in the 3 lines above; In this case, it is variable gain
           
```



## **Conditional Mutation**

Now, let's explore some advanced use of mutate and transumate; Here is 3 ways we can use for recode variable name.

Norice that `transumate` has the same functionality with `mutate`, but the only difference is that `transmuate` will give the related column, while `mutate` will give all the variables plus new variable you just created.

### **Method 1 ifelse**
```{r}
flight_subset %>% 
    transmute(arr_delay = arr_delay,
              status = ifelse(arr_delay > 0, "delayed", "On Time"))
```


### **Method 2 Case_when**

Here is another way we can write with the `ifelse` funciton. And this time, we can handle mutiple conditions

And this is every similar with if else statement in python, which makes the code very readable; Notice that the order for the ifelse statement matters, and `TRUE ~ "long"  ` means that whatever remained will be recode as `longer`.

```{r}
flight_distance <- flight_subset %>% 
    transmute(distance, 
              distance_type = case_when(
                  distance < 500 ~ "short", # if
                  distance < 1000 ~ "mid",  # elif
                  TRUE ~ "long"             # else
              ))

flight_distance
```

And the code above remind me of the `numpy` package in Python. In `numpy`, we have `numpy_where` function helps to vectorize and accelerate the processing speed; and it would be much quicker than `apply` function in python, and the format of code is eaxctly the same as above; please check out this [article for more information in python.](https://mp.weixin.qq.com/s?__biz=MzAxNTc0Mjg0Mg==&mid=2653296308&idx=1&sn=254bcaa023ac25e883a2b2b44293b3f3&chksm=802dd4a1b75a5db7d5b2955e3340eb6ab9095691b77fc8186f0a6f28b3e083779a6d8771ce5a&mpshare=1&scene=24&srcid=&sharer_sharetime=1578641340491&sharer_shareid=955e84dae4795ad4e079035b644897e4&key=e1c2dfa459d4b974c6dd352bb5e6fed432504266a86ed3fd459b09fa751cceec67cf1fb3264d3398971de102cf0d57692ff578d09e543bf3ca7907d569e30b148f293e478a9bf5550adfa2206159fcd9&ascene=14&uin=MTIxODU1NTA0NA%3D%3D&devicetype=Windows+10&version=6208006f&lang=zh_CN&exportkey=AS6hBBImDHRn3FsC9OovRLo%3D&pass_ticket=zyB9BMfXJt76KOucgp6FR47eJMSs9WN%2FGL8EM2wRAkmgb8cWCWPiZ7s9zbFo%2FePb)






### **Method 3 Recoding**


```{r}
flight_distance %>% mutate(distance_type = recode(distance_type,
            "long" = "long_distance", 
            "mid" = "mid_distance",
            "short" = "short _distance")) # here we do need necessarily double quote for long, mid, short
```




## `arrange`: changes the ordering of the rows

```{r}
flights %>% arrange(year, month, day)
flights %>% arrange(desc(dep_delay))
```


## `summarize` and `group_by` operations


In this case, we first goup all our data into different subgroup based on `tailnum`, and then do `summarize`; since we have nothing in the `summarize`, the code below just return the gorup.
```{r}
flight_subset %>% 
    group_by(tailnum) %>% 
    summarise()
```
































