---
title: General Introduction of Tidyverse Series1
author: ~
date: '2020-01-20'
slug: tidyverse-package-introduction-part1
categories: [R, Tidyverse]
tags: []
subtitle: 'Tidyverse package practice and Learning'
summary: 'An introduction of writting modern R code'
authors: [Ye Tao]
lastmod: '2020-01-20T16:46:19-08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


# **What is [tidyverse](https://www.tidyverse.org/)?**
The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures. This is my most favourite R package ecosystem for data manipulation, and is the reason that I still cannot completely get rid of using R.


-  To be more specific, It is a collection of R packages which are designed to be used together. 
  - `ggplot2`, for data visualisation
  - `dplyr`, for data manipulation
  - `tidyr`, for data tidying
  - `readr`, for data import
  - `purrr`, for functional programming
  - `tibble`, for tibbles, a modern re-imagining of data frames
  - `stringr`, for strings
  - `forcats`, for factors
  - and many other functions


# **Why tidyverse?**

There're a couple reasons:

- Base R functons are often slow and the implemenations sometimes are inconsistent.
- Writing tidyverse function use pip operator `%>%` is cleaner and more elegant.
- Lots of convenient functions for data manipulating and warggling helps accelerating data analysis process.



## Let's see five base functions given by dplyr

  - `select`: picks variables based on their names(Selecr columns). We do not need to add double quote on variable name
  - `filter`: picks cases based on their values; filter rows with some conditions.
  - `mutate`: adds new variables that are functions of existing variables
  - `arrange`: changes the ordering of the rows => reordering helps for data visualization
  - `summarize` or `summarise`: reduces multiple values down to a single summary
  - `group_by`: will group data into different categories, and we can add complicated calculation based on this function.
  
  
## **Obtain some data**

First of all, we need some data to work with.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13) # Please intstall this packages which contains some interating dataset.
flights %>% head
```


Data frame with columns information for `flight` dataset.

- `year, month, day`: Date of departure.

- `dep_time, arr_time`: Actual departure and arrival times (format HHMM or HMM), local tz.

- `sched_dep_time, sched_arr_time`: Scheduled departure and arrival times (format HHMM or HMM), local tz.

- `dep_delay, arr_delay`: Departure and arrival delays, in minutes. Negative times represent early departures/arrivals.

- `carrier`: Two letter carrier abbreviation. See airlines to get name.

- `flight`: Flight number.

- `tailnum`: Plane tail number. See planes for additional metadata.

- `origin, dest`: Origin and destination. See airports for additional metadata.

- `air_time`: Amount of time spent in the air, in minutes.

- `distance`: Distance between airports, in miles.

- `hour, minute`: Time of scheduled departure broken into hour and minutes.

- `time_hour`: Scheduled date and hour of the flight as a POSIXct date. Along with origin, can be used to join flights data to weather data.


## **Skimr Package**

This is a package used to quickly display summary statistics. `Skimr` handles different data types and returns a skim_df object which can be included in a tidyverse pipeline or displayed nicely for the human reader. It will shows you a distribution for each variable and outliers which marked with red color.

here I just use subset of the dataset for the purpose of demonstration. 


```{r}
library(skimr)
flight_subset <- flights %>% sample_n(5) # sample_frac, sample_integer
flight_subset %>% skim
```


## **Select functions**

You can select columns of the dataset by specifying the name in the `select` function
```{r, results='hide', error= FALSE}
flight_subset %>% 
    select(dep_time, arr_time)

flight_subset%>% 
    select(starts_with("arr_"))# specify the start word # ends_with()

flight_subset %>% 
    select(ends_with("_time")) # select all the columns with end word: time

flight_subset %>% 
    select(contains("dep")) # select the columns that contains "dep"

flight_subset %>% 
    select(-contains("dep")) # we can also put negation sign to select columns otherwise `dep`

# flight_subset %>% 
    # select(!contains("dep")) # we cannot use it sine this is not a logical vector, contains("dep")` must evaluate to column positions or names, not a logical vector

# Select use - for negation. While Filter use ! since it is logical operator

flight_subset %>% 
    select(dep_time: arr_time) # from dep_time to arr_time

flight_subset %>% 
    select(dep_time, everything()) # reordering , put the dep_time in the very front
```

- Notice that if you just need one single variable, you could use `pull` function. This `pull` function will give you a vector. While `select` function will give you a dataframe
- And you can use `rename` to rename your columns variable.

`pull` function directly gives the vector.
```{r}
flight_subset %>% 
    pull(dep_time) 
```


## **Filter Function**

```{r, results='hide'}
flight_subset %>% 
    filter(origin == "JFK") # filter the row with condition

flight_subset %>% 
    filter(distance > 1000)

# note that we are using a single `&` instead of `&&` as in base R
flight_subset %>% 
    filter(origin == "JFK" & distance > 100)

flight_subset %>% 
    filter(distance < 500 | distance > 1000) # this is or logical condition

flight_subset %>%
    filter(!between(distance, 500, 1000)) # since it is logical operator we use ! instead of !

# only keep the complete cases
flight_subset %>% 
    filter(complete.cases(flight_subset)) # This helps detect NA value in each row; use ! for negation

flight_subset %>% 
    slice(2:3) # Slice let you choose the rows depends on index.

flight_subset %>% 
    filter(carrier %in% c("DL", "UA"))
          #variable      value in columns we wanna pick
```

## **Chaining up function**

This is the my favourite part of tidyverse. We can use `%>%` to chain up functions instead of wrapping up lots of functions together. This helps making code seem more elegant and cleaner. Python also has the chainning ability.

```{r}
flight_subset %>% 
    select(origin, air_time) %>% 
    filter(origin == "JFK" & air_time > 10) %>% 
    rename(air_time_new = air_time)
```


## **Some more Example**

In this example we first select `origin` and `air_time`, and we filter out the row when `origin == "JFK"` and `air_time` < `mean(air_time)`. Here `mean(air_time)` is the mean of air_time in the flight_subset
```{r}
flight_subset %>% 
    select(origin, air_time) %>% 
    filter(origin == "JFK", air_time < mean(air_time, na.rm = TRUE )) #compare the mean with air_time in flight_subset
```

If it is difficult to understand, we can sperate above into two steps. And below gives the same result.

```{r}
# Step 1
mean_air_time = flight_subset %>% 
    pull(air_time) %>% 
    mean(na.rm = TRUE)

# Step 2
flight_subset %>% 
    select(origin, air_time) %>% 
    filter(origin == "JFK", air_time < mean_air_time) 

```


we can use `.` to represent the working data frame. These two below are the same
```{r}
flight_subset %>% 
    filter(complete.cases(.))

flight_subset %>% 
    filter(complete.cases(flight_subset))
```

## **Summary**
Just in case for avoiding cofusion; here is the summary

- x %>% f() is equivalent to f(x)
- x %>% f(.) is also equivalent to f(x)
- x %>% f(y) is equivalent to f(x, y)
- x %>% f(., y) is equivalent to f(x, y)

To have a concrete example, here we use `%>%` and `.` to compute binomial coefficients

```{r}
n <- 5
x <- 0:5
n %>% choose(x)

# or

x %>% choose(n, .) # . represent x, which is the vector we are working on

list(n=5, x = 0:5) %>% {
    choose(.$n, .$x)
}

# again . represents the thing we are working on with. Here is the list object => list(n-5, x=0:5), and let's call it A
# And then we pick out variable n and x from list A
```






